---
title: "Assignment 5 - Decision Trees"
author: "Charles Lang"
date: "November 9, 2016"
output: html_document
---
For this assignment we will be using data from the Assistments Intelligent Tutoring system. This system gives students hints based on how they perform on math problems. 

#Install & call libraries
```{r}
library(rpart)
library(party)
```

## Part I
```{r}
D1 <- read.csv("/Users/cici/assignment5-ysycici/intelligent_tutor.csv",header=TRUE)
```

##Classification Tree
First we will build a classification tree to predict which students ask a teacher for help, which start a new session, or which give up, based on whether or not the student completed a session (D1$complete) and whether or not they asked for hints (D1$hint.y). 
```{r}

c.tree <- rpart(action ~ hint.y + complete, method="class", data=D1) #Notice the standard R notion for a formula Y ~ X

#Look at the error of this tree
printcp(c.tree)

#Plot the tree
post(c.tree, file = "tree.ps", title = "Session Completion Action: 1 - Ask teacher, 2 - Start new session, 3 - Give up")

```
##### three numbers in each leaves are in accordance with the number of 1-ask teacher, 2-start new session, 3-give up. same number will always be the largest in one leaf, like: 2&128; 1&65; 3&76; 2&35; 3&44. others are different/wrong but less.


## Part II

#Regression Tree

We want to see if we can build a decision tree to help teachers decide which students to follow up with, based on students' performance in Assistments. We will create three groups ("teacher should intervene", "teacher should monitor student progress" and "no action") based on students' previous use of the system and how many hints they use. To do this we will be building a decision tree using the "party" package. The party package builds decision trees based on a set of statistical stopping rules.

#Take a look at our outcome variable "score"
```{r}
hist(D1$score)
```

#Create a categorical outcome variable based on student score to advise the teacher using an "ifelse" statement
```{r}
D1$advice <- ifelse(D1$score <=0.4, "intervene", ifelse(D1$score > 0.4 & D1$score <=0.8, "monitor", "no action"))
```

#Build a decision tree that predicts "advice" based on how many problems students have answered before, the percentage of those problems they got correct and how many hints they required
```{r}
score_ctree <- ctree(factor(advice) ~ prior_prob_count + prior_percent_correct + hints, D1)
```

#Plot tree
```{r}
plot(score_ctree)
```

Please interpret the tree, which two behaviors do you think the teacher should most closely pay attemtion to?

##### process: 1.check any coorelation between action and hints, prior_prob_count, prior_percent_correct respectively. If p<0.05 then reject the zero sum hypothesis (H0: there is no relationship). Choose the least one - hints (most significant one). Find the spilit point (0), then continue to split. 2.prior_prob_cont still has the coorelation, so continue to split (85 with lowest impurity). 3.continue to split hints (12 with lowest impurity), prior_percent_correct has relationship, split with 0.0629 (lowest impurity).

### the chart means the proprotion for the three actions. (add up together is 1) 

### two behaviors that the teacher should most closely pay attention to: 1.ask for hints >12  2.ask for hints >0 and <= 12, but has <=0.629 prior percent correct.


#Test Tree
Upload the data "intelligent_tutor_new.csv". This is a data set of a differnt sample of students doing the same problems in the same system. We can use the tree we built for the previous data set to try to predict the "advice" we should give the teacher about these new students. 

```{r}
#Upload new data

D2 <- read.csv("/Users/cici/assignment5-ysycici/intelligent_tutor_new.csv", header=TRUE)

#Generate predicted advice for new students based on tree generated from old students

D2$prediction <- predict(score_ctree, D2)

``` 
## Part III
Compare the predicted advice with the actual advice that these studnts recieved. What is the difference between the observed and predicted results?

### there are 84 differences out of 200 cases in D2.

```{r}
D2$advice <- ifelse(D2$score <=0.4, "intervene", ifelse(D2$score > 0.4 & D2$score <=0.8, "monitor", "no action"))
D2$diff <- ifelse(D2$advice==D2$prediction,0,1)
sum(D2$diff)
```

